|

### 2022-01-25 text/ManuscriptDraft.Rmd
Manuscript update

In this paper, we show that information on phenology at the level of individuals can be derived from image-based monitoring of flower phenology. We present and evaluate an automatic flower tracking and filtering algorithm. Many tracking algorithms, such as the Simple online and real-time tracking (SORT) [@sort], use Kalman filtering to predict the future location of an object by estimating its velocity [@kalmann]. However, Kalman filtering assumes linear motion of the object and does not handle abrupt motion well. As the flowers are constrained in movement by their stalk, their movement does not follow the assumption of linear motion. Some tracking algorithms, such as DeepSORT [@deepsort2017], take the tracking-by-detection approach in which convolutional neural networks (CNNs) are applied to leverage appearance information in order to continuously distinguish individuals through time, thus aiding the tracking algorithm. In our case, however, flowers are not clearly distinguishable and one flower may look less alike between two frames than two flowers in a single frame.

As the data extracted from time-lapse series of flower plots through season violate the assumptions of standard tracking mechanism, we base the track assignment on the minimum distance between centroids of bounding boxes and introduce three user defined parameters that can help optimise tracking performance.

@@

Several issues complicate the task of tracking individual flowers through the time-lapse series. First, monitoring the relatively slow process of flower development through full growing seasons (several months) requires low frame rates (several images per day). These frame rates capture the phenology of the flowers at very high detail, but fail to capture the full movement of flowers. For example, as the wind shifts, the flower heads changes direction, which in the image series happens instantaneously (i.e., between two consecutive frames). However, as they are constrained by their stalk, there is a limit to the distance they can move. With the tracking parameter **running mean**, we base the tracking on the distance between a point in the current frame and the running mean of the positions of a number of previous points in a track, taking advantage of the fact that flower move around a central point.

@@

The input data for the tracking algorithm are bounding box coordinates for each object in each frame given in the following format: frame number, x-min, y-min, x-max, y-max. Here, x-min, y-min is the top left corner of the bounding box and x-max, y-max is the bottom right corner. The bounding box centroid is calculated on the basis of these coordinates within the algorithm. Calculating tracking accuracy requires a ground truth ID for each bounding box which can be given in the sixth column.

As this data structure is outputted from both manual object annotations and object detection methods with CNNs, the method can be applied on both manually and automatically derived data.

@@

To explore the effect of the user parameters and to identify the optimal combination of parameters for our case of tracking flowers, we applied the tracking algorithm on each of the five image-series with every combination of a range of values for each parameter (3179 combinations): Values for max gap size and running mean were 0-160 with a step size of 10 and for max distance we used values 0-1000 with a step size of 100. Note that set to zero ignores the parameter altogether. We identified the setting(s) that returned the lowest number of track mismatches and compare the tracking results between optimal settings and all parameters set to zero.

@@

The optimal way of quantifying tracking performance depends on the goal of the tracking. To associate other information obtained in the images to the individual flower, for example flower visits, we want as much as possible of the track to be correct. To derive flowering length, in theory we just need to be able to connect the first and the last image of a flower and can ignore the intermediate points. Lastly, in some scenarios one may be interested in the number of flowers that existed in a plot, in which case we want the number of tracks obtained by automatic tracking to be as close as possible to the actual number of individuals in the series.

The multiple object tracking accuracy (MOTA) score quantifies tracking performance based on counts of tracking mismatches [@bernardin2008]. Mismatches occur when objects swap track identity because they are in close vicinity to each other or when an object periodically disappears and is assigned a new track identity when it reappears. Only the shifts in tracking identity are counted as mismatches while the number of points assigned to each track is not considered. We use three evaluation metrics: We calculate the number of mismatches and derive the MOTA score (number of mismatches in relation to the total number of points) to evaluate performance of our tracking and filtering algorithms. Finally, we compare the number of tracks identified by the automatic tracking with the true number of flowers in a series. These should ideally be equal.

@@

We identify a good value for eps by...

>>>>

In this paper, we show that information on phenology at the level of individuals can be derived from image-based monitoring of flower phenology. We present and evaluate an automatic flower tracking and filtering algorithm. Many tracking algorithms, such as the Simple online and real-time tracking (SORT) [@sort], use Kalman filtering to predict the future location of an object by estimating its velocity [@kalmann]. However, Kalman filtering assumes linear motion of the object and does not handle abrupt motion well. As the flowers are constrained in movement by their stalk, their movement does not follow the assumption of linear motion. Some tracking algorithms, such as DeepSORT [@deepsort2017], take the tracking-by-detection approach in which convolutional neural networks (CNNs) are applied to leverage appearance information in order to continuously distinguish individuals through time, thus aiding the tracking algorithm. In our case, however, flowers are not clearly distinguishable and one flower may look less alike between two frames than two flowers in a single frame. As the data extracted from time-lapse series of flower plots through season violate the assumptions of standard tracking mechanism, we base the track assignment on the minimum distance between centroids of bounding boxes and introduce three user defined parameters that can help optimise tracking performance.

@@

Several issues complicate the task of tracking individual flowers through time-lapse image-series. First, stand-alone monitoring of the relatively slow process of flower development through full growing seasons (several months) requires low frame rates (several images per day). These frame rates capture the phenology of the flowers at very high detail, but fail to capture the full movement of flowers. For example, as the wind shifts, the flower heads changes direction, which in the image series happens instantaneously (i.e., between two consecutive frames). However, as they are constrained by their stalk, there is a limit to the distance they can move. With the tracking parameter **running mean**, we base the tracking on the distance between a point in the current frame and the running mean of the positions of a number of previous points in a track, taking advantage of the fact that flower move around a central point.

@@

The input data for the tracking algorithm are bounding box coordinates for each object in each frame given in the following format: frame number, x-min, y-min, x-max, y-max. Here, x-min, y-min is the top left corner of the bounding box and x-max, y-max is the bottom right corner. The bounding box centroid is calculated on the basis of these coordinates within the algorithm. Calculating tracking accuracy requires a ground truth ID for each bounding box which can be given as the sixth column. As this data structure is outputted from both manual object annotations and object detection methods using CNNs, the method can be applied on both manually and automatically derived data. Further, as the algorithm handles frames sequentially, it can be applied both offline and online.

@@

To explore the effect of the user parameters and to identify the optimal combination of parameters for our case of tracking flowers, we applied the tracking algorithm on each of the five image-series with every combination of a range of values for each parameter (3179 combinations): Values for max gap size and running mean were 0-160 with a step size of 10 and for max distance we used values 0-1000 with a step size of 100. Note that max distance set to zero ignores the parameter altogether. We identified the setting(s) that returned the lowest number of track mismatches and compare the tracking results between optimal settings and all parameters set to zero.

@@

The optimal way of quantifying tracking performance depends on the goal of the tracking. To associate other information obtained in the images to the individual flower, for example flower visits, requires as much as possible of the track to be correct. To derive flowering length, in theory just requires correct association between the first and the last image of a flower while intermediate points can be ignored. Lastly, in some scenarios one may be interested in the number of flowers that existed in a plot, in which case the number of tracks obtained by automatic tracking should closely resemble the actual number of individuals in the series.

The commonly used multiple object tracking accuracy (MOTA) score quantifies tracking performance based on counts of tracking mismatches [@bernardin2008]. Mismatches occur when objects swap track identity because they are in close vicinity to each other or when an object periodically disappears and is assigned a new track identity when it reappears. Only the shifts in tracking identity are counted as mismatches while the number of points assigned to each track is not considered. We use three evaluation metrics: We calculate the number of mismatches and derive the MOTA score (number of mismatches in relation to the total number of points) to evaluate performance of our tracking and filtering algorithms. Finally, we compare the number of tracks identified by the automatic tracking with the true number of flowers in a series. These should ideally be equal. However, a correct number of tracks (i.e. number of tracks equal to number of objects) does not necessarily translate to correct tracking, since individual tracks can be erroneously split by the tracking algorithm and several tracks can be combined. We present the tracking performance by both the lowets number of mismatches/highest MOTA given overall and for combinations that returned a number opf tracks equal to the number of flowers in the series.

@@
